---
title: DANL Project
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - Henry Socash 
  - Alban Kuka
  - Victoria Pavlock
  - Jackson Boyer
  - Jeremy Sun

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

About this project :clap:
-we will be observing how Magnitude varies by Depth, Location, and Time of day for the earthquakes recorded in 2023 

```{r}
earthquake <- read_csv('C:/Users/hsoca/OneDrive/Desktop/earthquake data.csv')
```


# Data

this data set, according to its Kaggle Description, which can be found at https://www.kaggle.com/datasets/syedanwarafridi/earthquake-events-worldwide-2023?rvi=1 is "a comprehensive catalog capturing seismic activity from January 1, 2023, to December 31, 2023. This dataset provides a detailed record of earthquakes globally, including crucial information such as date and time, geographic coordinates, depth, magnitude, and the affected region's land type and country"
```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(earthquake)
```

## Summary Statistics

 Here are some summary statistics of the numeric variables in this dataset. 
Depth is measured in Kilometers and Magnitude is measured on the Richter
```{r}
eq1 <- earthquake %>% select(Magnitude,Depth)
skim(eq1)
```


## Exploratory Data Analysis 

 here shows top ten regions and countries that experienced the most earthquakes in 2023
```{r}
quaked <- earthquake %>% 
  group_by(Country , Lands) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=10) %>% 
  head(quaked, n = 10)
```
 as we can see Turkey is head and shoulders above other countries with three separate regions of the same country all making the top 10 most quaked country/region
 generally we can see that the most common places for earthquakes to occur are around the Mediterranean and south american coast 


## more exploratory data analysis

 here is a data set showing the months where the most earthquakes took place 
```{r}
earthquake$Month <- rowSums(earthquake[,4:5])
earthquake$Month[1:6055] <- "NOVEMBER"
earthquake$Month[6056:14543] <- "OCTOBER"
earthquake$Month[14544:23404] <- "SEPTEMBER"
earthquake$Month[23405:31631] <- "AUGUST"
earthquake$Month[31632:40673] <- "JULY"
earthquake$Month[40674:48647] <- "JUNE"
earthquake$Month[48648:53539] <- "MAY"

densemonth <- earthquake %>% 
  group_by(Month) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=7)

```

## proportion of earthquakes per month, per country 

```{r}
dense <- earthquake %>% 
  group_by(Month,Country,Lands) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=7) %>% 
  head(dense, n=15)
ggplot(dense,
       aes(y = Month,
           fill = Country)) +
  geom_bar(show.legend = T)
```
we can see that most countries sporadically receive earthquakes around the year, whereas Turkeys threat of an earthquake remains constant year round.

## why is this important?

this data can be helpful for anyone wishing to visit some of the more frequently earth quaked countries, wanting to know when the safest time of year to vacation is.
Understanding where and when earthquakes appear can also be helpful for understanding more about the causes of earthquakes and possibly how they can be avoided.
