---
title: DANL Project
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - Henry Socash 
  - Name 2
  - Name 3
  - Name 4
  - Name 5

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

About this project :clap:
-we will be observing how Magnitude varies by Depth, Location, and Time of day for the earthquakes recorded in 2023 

```{r}
eq <- read_csv('C:/Users/hsoca/OneDrive/Desktop/earthquake data.csv')
```


# Data

this data set contains information about all earthquakes that occurred in 2023, displaying magnitude, location, depth, and time of the event. 
```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(eq)
```

## Summary Statistics

Here are some summary statistics of the numeric variables in this dataset. 
Depth is measured in Kilometers and Magnitude is measured on the Richter
```{r}
eq1 <- eq %>% select(Magnitude,Depth)
skim(eq1)
```


## Exploratory Data Analysis 

here shows top ten regions and countries that experienced the most eathquakes in 2023
```{r}
quaked <- earthquake %>% 
  group_by(Country , Lands) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=10) %>% 
  head(quaked, n = 10)
```
## more exploratory data analysis

here is a data set showing the months where the most earthquakes took place 
```{r}
earthquake$Month <- rowSums(earthquake[,4:5])
earthquake$Month[1:6055] <- "NOVEMBER"
earthquake$Month[6056:14543] <- "OCTOBER"
earthquake$Month[14544:23404] <- "SEPTEMBER"
earthquake$Month[23405:31631] <- "AUGUST"
earthquake$Month[31632:40673] <- "JULY"
earthquake$Month[40674:48647] <- "JUNE"
earthquake$Month[48648:53539] <- "MAY"

densemonth <- earthquake %>% 
  group_by(Month) %>% 
  summarize(count =n()) %>% 
  arrange(desc(count)) %>% 
  filter(row_number() <=7)

```

## MPG and a Type of Cars
The following boxplot shows how the distribution of highway MPG (`hwy`) varies by a type of cars (`class`) :blue_car: :truck: :minibus:.


```{r}

```


